Generative Models for scRNA

### 2018-04-02 generativemodelsforscrna/outline.tex
consistency proof, such as it is

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider the following figure:

\includegraphics[width=0.8\textwidth]{pics/Z1gn}

@@

Another way to see that the model isn't quite right is to look at posterior predictive p values.  That is, for each entry of the matrix, we can compute the cumulative distribution function of that entry in light of the parameters predicted by the posterior mode.  Here we have done that for the first 1000 entries.  The results are not exactly encouraging:

\includegraphics[width=0.8\textwidth]{pics/resids}

We see that for some genes the $p$ values (which should be distributed uniformly) are almost always near 1.  We can dig deeper into this by looking at individual genes, and comparing the expected value for that gene under the model with the actual value.  There are clearly some systematic differences:

\includegraphics[width=0.8\textwidth]{pics/residsdetail}

One potential reason for this systematic error might be that the negative binomial data distribution is inappropriate.  To diagnose this, we fitted a zero-inflated negative binomial distribution to each collection of cells Allen Institute dubbed a ``cluster.''  We then measured the goodness-of-fit, in terms of whether various moments matched.  In particular, we compared the empirical mean, standard deviation, and proportion of zeros to the corresponding statistics which would be predicted by the learned models.

\includegraphics[width=0.8\textwidth]{pics/zinbfitmatch}

@@

Another possible source of trouble is that in the particular model we use we have made certain assumption about how $\alpha,\beta,h$ can vary.  This could be a big problem.  To see why, we looked at each the fitted parameters of the zero-inflated negative binomial distribution for each cell-type.  This yields a kind of three-dimensional scatterplot, in which each dot is a cell type, and the location indicates the value of each of its three fitted parameters.  Here we show three projections of that scatterplot:

\includegraphics[width=0.8\textwidth]{pics/zinbfitdiversity}

>>>>

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider Figure \ref{fig:Z1gn}:

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/Z1gn}
\caption{\label{fig:Z1gn}}
\end{figure}

@@

Another way to see that the model isn't quite right is to look at posterior predictive p values.  That is, for each entry of the matrix, we can compute the cumulative distribution function of that entry in light of the parameters predicted by the posterior mode.  In Figure \ref{fig:resids} we have done that for the first 1000 entries.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/resids}
\caption{\label{fig:resids}}
\end{figure}

These results are actually fairly encouraging.  Mostly it looks basically like white noise.  However, we do see that for some genes the $p$ values are almost always near 1: these are genes which express essentially never, and so have a high chance of turning off, and so their $p$ value is nearly $1$ whenever they are off, which is virtually always.  If we restrict ourselves to those genes whose average probability of going off is less than 20\%, we obtain Figure \ref{fig:residslesszeros}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/residslesszeros}
\caption{\label{fig:residslesszeros}}
\end{figure}

Another way to dig deeper into the underfitting question is by looking at individual genes, and comparing the expected value for that gene under the model with the actual value.  There are clearly some systematic differences, as we see in Figure \ref{fig:residsdetail}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/residsdetail}
\caption{\label{fig:residsdetail}}
\end{figure}

One potential reason for this systematic error might be that the negative binomial data distribution is inappropriate.  To diagnose this, we fitted a zero-inflated negative binomial distribution to each collection of cells Allen Institute dubbed a ``cluster.''  We then measured the goodness-of-fit, in terms of whether various moments matched.  In particular, we compared the empirical mean, standard deviation, and proportion of zeros to the corresponding statistics which would be predicted by the learned models.  We show this in Figure \ref{fig:zinbfitmatch}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/zinbfitmatch}
\caption{\label{fig:zinbfitmatch}}
\end{figure}

@@

Another possible source of trouble is that in the particular model we use we have made certain assumption about how $\alpha,\beta,h$ can vary.  This could be a big problem.  To see why, we looked at each the fitted parameters of the zero-inflated negative binomial distribution for each cell-type.  This yields a kind of three-dimensional scatterplot, in which each dot is a cell type, and the location indicates the value of each of its three fitted parameters.  Here we show three projections of that scatterplot in Figure \ref{fig:zinbfitdiversity}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/zinbfitdiversity}
\caption{\label{fig:zinbfitdiversity}}
\end{figure}

### 2018-04-02 generativemodelsforscrna/paper.tex
consistency proof, such as it is

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider Figure \ref{fig:Z1gn}:

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/Z1gn}
\caption{\label{fig:Z1gn}}
\end{figure}

@@

Another way to see that the model isn't quite right is to look at posterior predictive p values.  That is, for each entry of the matrix, we can compute the cumulative distribution function of that entry in light of the parameters predicted by the posterior mode.  In Figure \ref{fig:resids} we have done that for the first 1000 entries.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/resids}
\caption{\label{fig:resids}}
\end{figure}

These results are actually fairly encouraging.  Mostly it looks basically like white noise.  However, we do see that for some genes the $p$ values are almost always near 1: these are genes which express essentially never, and so have a high chance of turning off, and so their $p$ value is nearly $1$ whenever they are off, which is virtually always.  If we restrict ourselves to those genes whose average probability of going off is less than 20\%, we obtain Figure \ref{fig:residslesszeros}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/residslesszeros}
\caption{\label{fig:residslesszeros}}
\end{figure}

Another way to dig deeper into the underfitting question is by looking at individual genes, and comparing the expected value for that gene under the model with the actual value.  There are clearly some systematic differences, as we see in Figure \ref{fig:residsdetail}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/residsdetail}
\caption{\label{fig:residsdetail}}
\end{figure}

One potential reason for this systematic error might be that the negative binomial data distribution is inappropriate.  To diagnose this, we fitted a zero-inflated negative binomial distribution to each collection of cells Allen Institute dubbed a ``cluster.''  We then measured the goodness-of-fit, in terms of whether various moments matched.  In particular, we compared the empirical mean, standard deviation, and proportion of zeros to the corresponding statistics which would be predicted by the learned models.  We show this in Figure \ref{fig:zinbfitmatch}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/zinbfitmatch}
\caption{\label{fig:zinbfitmatch}}
\end{figure}

@@

Another possible source of trouble is that in the particular model we use we have made certain assumption about how $\alpha,\beta,h$ can vary.  This could be a big problem.  To see why, we looked at each the fitted parameters of the zero-inflated negative binomial distribution for each cell-type.  This yields a kind of three-dimensional scatterplot, in which each dot is a cell type, and the location indicates the value of each of its three fitted parameters.  Here we show three projections of that scatterplot in Figure \ref{fig:zinbfitdiversity}.

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/zinbfitdiversity}
\caption{\label{fig:zinbfitdiversity}}
\end{figure}

>>>>

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider the following figure:

\includegraphics[width=0.8\textwidth]{pics/Z1gn}

@@

Another way to see that the model isn't quite right is to look at posterior predictive p values.  That is, for each entry of the matrix, we can compute the cumulative distribution function of that entry in light of the parameters predicted by the posterior mode.  Here we have done that for the first 1000 entries.

\includegraphics[width=0.8\textwidth]{pics/resids}

These results are actually fairly encouraging.  Mostly it looks basically like white noise.  However, we do see that for some genes the $p$ values are almost always near 1: these are genes which express essentially never, and so have a high chance of turning off, and so their $p$ value is nearly $1$ whenever they are off, which is virtually always.

Anotehr way to dig deeper into the underfitting question is by looking at individual genes, and comparing the expected value for that gene under the model with the actual value.  There are clearly some systematic differences:

\includegraphics[width=0.8\textwidth]{pics/residsdetail}

One potential reason for this systematic error might be that the negative binomial data distribution is inappropriate.  To diagnose this, we fitted a zero-inflated negative binomial distribution to each collection of cells Allen Institute dubbed a ``cluster.''  We then measured the goodness-of-fit, in terms of whether various moments matched.  In particular, we compared the empirical mean, standard deviation, and proportion of zeros to the corresponding statistics which would be predicted by the learned models.

\includegraphics[width=0.8\textwidth]{pics/zinbfitmatch}

@@

Another possible source of trouble is that in the particular model we use we have made certain assumption about how $\alpha,\beta,h$ can vary.  This could be a big problem.  To see why, we looked at each the fitted parameters of the zero-inflated negative binomial distribution for each cell-type.  This yields a kind of three-dimensional scatterplot, in which each dot is a cell type, and the location indicates the value of each of its three fitted parameters.  Here we show three projections of that scatterplot:

\includegraphics[width=0.8\textwidth]{pics/zinbfitdiversity}
